### 2021-09-06 manuscript/methodology.tex
until "Convolutional equivalent layer"

Equation \ref{eq:integral-sum_mag} can be rewritten in matrix notation as follows:
\begin{equation}
\mathbf{d}(\mathbf{p}) = \mathbf{A} \mathbf{p} \: ,
\label{eq:predicted-data-vector_mag}
\end{equation}
where $\mathbf{d}(\mathbf{p})$ is the $N \times 1$ predicted data vector with $i$th element defined
be the predicted total-field anomaly $\Delta T_{i}$ (equation \ref{eq:integral-sum_mag}),
$\mathbf{p}$ is the $M \times 1$ parameter vector whose $j$th element is the magnetic moment intensity
$p_{j}$ of the $j$th dipole and $\mathbf{A}$ is the $N \times M$ sensitivity matrix with element 
$ij$ defined by the harmonic function $a_{ij}$ (equation \ref{eq:aij_mag}).
In the classical equivalent-layer technique, the common approach for 
estimating the parameter vector $\mathbf{p}$ from the observed 
total-field anomaly data $\mathbf{d}^{o}$ is solving the least-squares normal equations
\begin{equation}
\mathbf{A}^{\top}\mathbf{A} \: \mathbf{p} = 
\mathbf{A}^{\top} \mathbf{d}^{o} \: .
\label{eq:normal-equations}
\end{equation}
This equation is usually solved by first computing the Cholesky factor of matrix
$\mathbf{A}^{\top}\mathbf{A}$ and then using it to solve the linear system
\citep[][ p. 262]{golub-vanloan2013}. 
The estimated parameter vector obtained by following this approach will be 
referenced throughout this work as the \textit{classical solution}.

@@

By using equations \ref{eq:hxx_regular}--\ref{eq:1_rij_regular} to compute 
$a^{\alpha\beta}_{ij}$ (equation \ref{eq:aij_alpha_beta}), we can show that 
matrices $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ (equation \ref{eq:A_expand}) assume
well-defined structures that can be conveniently
represented with \textit{block indices} $q$ and $p$ \citep{takahashi2020convolutional}.
These indices are defined by the integer functions $\Delta k_{ij}$ and $\Delta l_{ij}$ 
(equations \ref{eq:Delta_kij} and \ref{eq:Delta_lij}), in terms of the indices $i$ 
of the observation points $(x_{i}, y_{i}, z_{0})$ and $j$ of the equivalent sources
$(x_{j}, y_{j}, z_{c})$.
For $x$-\textit{oriented grids} (Figure \ref{fig:regular-grids}), $Q = N_{y}$, $P = N_{x}$ 
and the block indices $q$ and $p$ are given by:
\begin{equation}
q \equiv q(i, j) = \Delta l_{ij}
\label{eq:q-x-oriented}
\end{equation}
and
\begin{equation}
p \equiv p(i, j) = \Delta k_{ij} \: ,
\label{eq:p-x-oriented}
\end{equation}
where $\Delta k_{ij}$ and $\Delta l_{ij}$ (equations \ref{eq:Delta_kij} and \ref{eq:Delta_lij}) 
are defined by integer functions $k(i)$, $k(j)$, $l(i)$ and $l(j)$ given by equations 
\ref{eq:k-x-oriented} and \ref{eq:l-x-oriented}.
For $y$-\textit{oriented grids} (Figure \ref{fig:regular-grids}), $Q = N_{x}$, $P = N_{y}$ and 
the block indices $q$ and $p$ are given by:
\begin{equation}
q \equiv q(i, j) = \Delta k_{ij}
\label{eq:q-y-oriented}
\end{equation}
and
\begin{equation}
p \equiv p(i, j) = \Delta l_{ij} \: ,
\label{eq:p-y-oriented}
\end{equation}
where $\Delta k_{ij}$ and $\Delta l_{ij}$ (equations \ref{eq:Delta_kij} and \ref{eq:Delta_lij}) 
are defined by integer functions $k(i)$, $k(j)$, $l(i)$ and $l(j)$ given by equations 
\ref{eq:k-y-oriented} and \ref{eq:l-y-oriented}.
Equations \ref{eq:q-x-oriented}--\ref{eq:p-y-oriented} differ from those presented by
\citet{takahashi2020convolutional} in the absence of the module.

@@

Note that, for $x$-oriented grids, matrices $\Delta\mathbf{K}$ (equation \ref{eq:DK-matrix-x-oriented})
and $\Delta\mathbf{L}$ (equation \ref{eq:DL-matrix-x-oriented}) define the block indices
$p$ (equation \ref{eq:p-x-oriented}) and $q$ (equation \ref{eq:q-x-oriented}), respectively.
In this case, they are composed of $Q \times Q$ blocks with $P \times P$ elements each, where 
$Q = N_{y}$ and $P = N_{x}$. 
For $y$-oriented grids, matrices $\Delta\mathbf{K}$ (equation \ref{eq:DK-matrix-y-oriented})
and $\Delta\mathbf{L}$ (equation \ref{eq:DL-matrix-y-oriented}) define the block indices
$q$ (equation \ref{eq:q-y-oriented}) and $p$ (equation \ref{eq:p-y-oriented}), respectively.
In this case, they are also composed of $Q \times Q$ blocks with $P \times P$ elements each, 
but now $Q = N_{x}$ and $P = N_{y}$.
The examples shown by equations \ref{eq:DK-matrix-x-oriented}--\ref{eq:DL-matrix-y-oriented}
also illustrate that, regardless of grid orientation, (i) the block index $q$ is constant 
inside each block; (ii) blocks disposed along the same block diagonal are equal to each other; 
(iii) the block index $p$ is constant on each diagonal of a given block; 
(iv) elements of a given block located on the same diagonal are also equal do each other.
The results obtained with the small grid shown in Figure \ref{fig:regular-grids}
can be easily generalized for larger grids.
Based on the well-defined structure of block indices, we can define 
matrices $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ in a general form
\begin{equation}
\mathbf{A}_{\boldsymbol{\alpha\beta}} = \begin{bmatrix}
\mathbf{A}_{\boldsymbol{\alpha\beta}}^{0}   & \mathbf{A}_{\boldsymbol{\alpha\beta}}^{-1} & \cdots          & \mathbf{A}_{\boldsymbol{\alpha\beta}}^{-Q+1} \\
\mathbf{A}_{\boldsymbol{\alpha\beta}}^{1}   & \ddots          & \ddots          & \vdots           \\ 
\vdots           & \ddots          & \ddots          & \mathbf{A}_{\boldsymbol{\alpha\beta}}^{-1}   \\
\mathbf{A}_{\boldsymbol{\alpha\beta}}^{Q-1} & \cdots          & \mathbf{A}_{\boldsymbol{\alpha\beta}}^{1}  & \mathbf{A}_{\boldsymbol{\alpha\beta}}^{0}
\end{bmatrix}_{N \times N} \: ,
\label{eq:BTTB_A_alpha_beta}
\end{equation}
with blocks $\mathbf{A}_{\boldsymbol{\alpha\beta}}^{q}$ given by
\begin{equation}
\mathbf{A}_{\boldsymbol{\alpha\beta}}^{q} = \begin{bmatrix}
a^{\alpha\beta}_{q0}   & a^{\alpha\beta}_{q(-1)} & \cdots  & a^{\alpha\beta}_{q(-P+1)} \\
a^{\alpha\beta}_{q1}   & \ddots     & \ddots  & \vdots       \\ 
\vdots      & \ddots     & \ddots  & a^{\alpha\beta}_{q(-1)}   \\
a^{\alpha\beta}_{q(P-1)} & \cdots     & a^{\alpha\beta}_{q1}  & a^{\alpha\beta}_{q0}
\end{bmatrix}_{P \times P} \: .
\label{eq:Aq_block}
\end{equation}
This well-defined structure (equations \ref{eq:BTTB_A_alpha_beta} and \ref{eq:Aq_block}) 
of matrix components $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ 
(equation \ref{eq:A_expand}) is called Block-Toeplitz Toeplitz-Block (BTTB) 
\citep[e.g., ][ p. 67]{chan-jin2007}.

@@

Let $\mathbf{A}_{\boldsymbol{xz}}$ be a matrix component with elements
$a^{xz}_{ij}$ (equation \ref{eq:aij_alpha_beta}) defined by the second derivative
$h^{xz}_{ij}$ (equation \ref{eq:hxz_regular}). It can be easily verified from equations
\ref{eq:Delta_kij_symmetry}--\ref{eq:1_rij_symmetry} that $h^{xz}_{ij} = -h^{xz}_{ji}$.
As a consequence, $a^{xz}_{ij} = -a^{xz}_{ji}$, which means that 
\begin{equation}
\mathbf{A}_{\boldsymbol{xz}} = -\left( \mathbf{A}_{\boldsymbol{xz}} \right)
\label{eq:Axz_symmetry}
\end{equation} 
for any grid orientation.
For $x$-oriented grids (Figure \ref{fig:regular-grids}), the block indices $q$ and $p$ are defined 
by equations \ref{eq:q-x-oriented} and 
\ref{eq:p-x-oriented} and $a^{xz}_{qp}$ can be rewritten as follows:
\begin{equation}
a^{xz}_{qp} = c_{m} \, \frac{\mu_{0}}{4\pi} 
\left( F_{x} u_{z} + F_{z} u_{x} \right) \frac{3 \left( p \, \Delta_{x} \right) \Delta_{z}}{r_{qp}^{5}}
\: ,
\label{eq:aqp_xz_x_oriented}
\end{equation}
with $\tfrac{1}{r_{qp}}$ defined by equation \ref{eq:1_rqp_x_oriented}.
In this case, we can see that
\begin{equation}
\mathbf{A}_{\boldsymbol{xz}}^{q} = \mathbf{A}_{\boldsymbol{xz}}^{(-q)}
\label{eq:Axz_q_external_block_symmetry_x_oriented}
\end{equation}
and 
\begin{equation}
\mathbf{A}_{\boldsymbol{xz}}^{q} = -\left( \mathbf{A}_{\boldsymbol{xz}}^{q} \right)^{\top} \: .
\label{eq:Axz_q_internal_block_symmetry_x_oriented}
\end{equation}
This structure is called \textit{symmetric-Block-Toeplitz skew symmetric-Toeplitz-Block} and is 
valid only for $x$-oriented grids.
For $y$-oriented grids (Figure \ref{fig:regular-grids}), the block indices $q$ and $p$ are 
defined by equations \ref{eq:q-y-oriented} and 
\ref{eq:p-y-oriented} and $a^{xz}_{qp}$ can be rewritten as follows:
\begin{equation}
a^{xz}_{qp} = c_{m} \, \frac{\mu_{0}}{4\pi} 
\left( F_{x} u_{z} + F_{z} u_{x} \right) \frac{3 \left( q \, \Delta_{x} \right) \Delta_{z}}{r_{qp}^{5}} \: ,
\label{eq:aqp_xz_y_oriented}
\end{equation}
with $\tfrac{1}{r_{qp}}$ defined by equation \ref{eq:1_rqp_y_oriented}.
Now, we conclude that
\begin{equation}
\mathbf{A}_{\boldsymbol{xz}}^{q} = -\mathbf{A}_{\boldsymbol{xz}}^{(-q)}
\label{eq:Axz_q_external_block_symmetry_y_oriented}
\end{equation}
and 
\begin{equation}
\mathbf{A}_{\boldsymbol{xz}}^{q} = \left( \mathbf{A}_{\boldsymbol{xz}}^{q} \right)^{\top} \: .
\label{eq:Axz_q_internal_block_symmetry_y_oriented}
\end{equation}
This structure is called \textit{skew symmetric-Block-Toeplitz symmetric-Toeplitz-Block} and is 
valid only for $y$-oriented grids.

The same reasoning can be followed to show that
\begin{equation}
\mathbf{A}_{\boldsymbol{yz}} = -\left( \mathbf{A}_{\boldsymbol{yz}} \right)
\label{eq:Ayz_symmetry}
\end{equation} 
for any grid orientation. Besides, we can also show that
\begin{equation}
\mathbf{A}_{\boldsymbol{yz}}^{q} = -\mathbf{A}_{\boldsymbol{yz}}^{(-q)}
\label{eq:Ayz_q_external_block_symmetry_x_oriented}
\end{equation}
and 
\begin{equation}
\mathbf{A}_{\boldsymbol{yz}}^{q} = \left( \mathbf{A}_{\boldsymbol{yz}}^{q} \right)^{\top}
\label{eq:Ayz_q_internal_block_symmetry_x_oriented}
\end{equation}
for $x$-oriented grids, while
\begin{equation}
\mathbf{A}_{\boldsymbol{yz}}^{q} = \mathbf{A}_{\boldsymbol{yz}}^{(-q)}
\label{eq:Ayz_q_external_block_symmetry_y_oriented}
\end{equation}
and 
\begin{equation}
\mathbf{A}_{\boldsymbol{yz}}^{q} = -\left( \mathbf{A}_{\boldsymbol{yz}}^{q} \right)^{\top}
\label{eq:Ayz_q_internal_block_symmetry_y_oriented}
\end{equation}
for $y$-oriented grids.

@@

%======================================================================================
\subsection{Standard Conjugate Gradient Least Squares (CGLS) method}
%======================================================================================

The computational cost associated with the classical method for estimating the parameter 
vector $\mathbf{p}$ by solving the linear system \ref{eq:normal-equations} can be very high 
or ever prohibitive when dealing with large data sets. In these cases, a well-known alternative
is solving the normal equations (equation \ref{eq:normal-equations}) iteratively by 
using the standard Conjugate Gradient Least Squares (CGLS) method. Below we present a pseudocode 
for the standard CGLS method:

@@

Setting a convergence criteria based on the minimum tolerance of the residuals is a good 
option to carry out this algorithm efficiently and still obtaining very good results. 
Another possibility is to set an invariance to the Euclidean norm of residuals between 
iterations, which would increase algorithm runtime, but with smaller residuals. 
We chose the first option, as we achieve satisfactory results. 
The estimated parameter vector obtained by using the standard CGLS method 
(Algorithm \ref{al:std-cgls-algorithm}) will be referenced throughout this work as 
the \textit{standard CGLS solution}.

Note that the standard CGLS solution (Algorithm \ref{al:std-cgls-algorithm}) requires 
neither inverse matrix nor matrix-matrix product. Instead, it only requires: one matrix-vector 
product out of the loop and two matrix-vector products per iteration (in steps 3 and 6). 
In the following subsections, we will show how to compute these three matrix-vector products 
efficiently by exploring the structure of the sensitivity matrix $\mathbf{A}$.

@@

A new auxuliary linear system is constructed to carry the matrix-vector product:
\begin{equation}
\mathbf{w} = \mathbf{C} \mathbf{v} \: ,
\label{eq:w_Cv}
\end{equation}
where
\begin{equation}
\mathbf{w} = \begin{bmatrix}
\mathbf{w}_{0} \\
\vdots \\
\mathbf{w}_{Q - 1} \\
\mathbf{0}_{2N \times 1}
\end{bmatrix}_{4N \times 1} \quad ,
\label{eq:w-vector}
\end{equation}
\begin{equation}
\mathbf{w}_{q} = \begin{bmatrix}
\mathbf{d}_{q}(\mathbf{p}) \\
\mathbf{0}_{P \times 1}
\end{bmatrix}_{2P \times 1}
\label{eq:wq-vector} \quad ,
\end{equation}
\begin{equation}
\mathbf{v} = \begin{bmatrix}
\mathbf{v}_{0} \\
\vdots \\
\mathbf{v}_{Q - 1} \\
\mathbf{0}_{2N \times 1}
\end{bmatrix}_{4N \times 1} \quad ,
\label{eq:v-vector}
\end{equation}
and
\begin{equation}
\mathbf{v}_{q} = \begin{bmatrix}
\mathbf{p}_{q} \\
\mathbf{0}_{P \times 1}
\end{bmatrix}_{2P \times 1}
\label{eq:vq-vector} \quad ,
\end{equation}
where $\mathbf{C}$ (equation \ref{eq:w_Cv}) is a $4N \times 4N$ non-symmetric (BCCB) resulted from transforming $\mathbf{A}$ (equation \ref{eq:aij_mag}). Without having to calculate the whole BCCB matrix, its first column can be used to carry the multiplication of this new system (equation \ref{eq:w_Cv}). Appendix A and C in \cite{takahashi2020convolutional} shows how to use the 2D-FFT to compute the eigenvalues of matrix $\mathbf{C}$, store in a $2Q \times 2P$ matrix using the $vec$-operator and to carry the matrix-vector product. Denoting the matrix $\mathbf{L}$ as the eigenvalues matrix follows:
\begin{equation}
	\mathbf{F}_{2Q}^{\ast} \left[ 
	\mathbf{L} \circ \left(\mathbf{F}_{2Q} \, \mathbf{V} \, \mathbf{F}_{2P} \right) 
	\right] \mathbf{F}_{2P}^{\ast} = \mathbf{W} \: ,
	\label{eq:DFT-system}
\end{equation}
where the symbol ``$\circ$'' references the Hadamard product, i.e., a element-wise complex multiplication between the eingenvalues and the 2D-FFT of the matrix rearranged along the rows of the parameters $\mathbf{V}$ (equation \ref{eq:v-vector}) using the same $vec$-operator. The resulting inverse 2D-FFT denoted by $\mathbf{F}_{2Q}^{\ast}  \otimes \mathbf{F}_{2P}^{\ast}$ is also a $2Q \times 2P$ matrix ($\mathbf{W}$) that can be rearranged to the predicted data vector $\mathbf{d}(\hat{\mathbf{p}})$ size $N$.

>>>>

Equation \ref{eq:integral-sum_mag} can be rewritten in matrix notation as follows:
\begin{equation}
\mathbf{d}(\mathbf{p}) = \mathbf{A} \mathbf{p} \: ,
\label{eq:predicted-data-vector_mag}
\end{equation}
where $\mathbf{d}(\mathbf{p})$ is the $N \times 1$ predicted data vector with $i$th element defined
be the predicted total-field anomaly $\Delta T_{i}$ (equation \ref{eq:integral-sum_mag}),
$\mathbf{p}$ is the $M \times 1$ parameter vector whose $j$th element is the magnetic moment intensity
$p_{j}$ of the $j$th dipole and $\mathbf{A}$ is the $N \times M$ sensitivity matrix with element 
$ij$ defined by the harmonic function $a_{ij}$ (equation \ref{eq:aij_mag}).
In the classical equivalent-layer technique, the common approach for 
estimating the parameter vector $\mathbf{p}$ from the observed 
total-field anomaly data $\mathbf{d}^{o}$ is solving the least-squares normal equations
\begin{equation}
\mathbf{A}^{\top}\mathbf{A} \: \mathbf{p} = 
\mathbf{A}^{\top} \mathbf{d}^{o} \: .
\label{eq:normal-equations}
\end{equation}
Equation \ref{eq:normal-equations} is usually solved by first computing the Cholesky 
factor of matrix $\mathbf{A}^{\top}\mathbf{A}$ and then using it to solve the linear 
system \citep[][ p. 262]{golub-vanloan2013}. 
This approach to estimate the parameter vector will be 
referenced throughout this work as the \textit{classical method}.

@@

By using equations \ref{eq:hxx_regular}--\ref{eq:1_rij_regular} to compute 
$a^{\alpha\beta}_{ij}$ (equation \ref{eq:aij_alpha_beta}), we can show that 
matrices $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ (equation \ref{eq:A_expand}) assume
well-defined structures that can be conveniently
represented with \textit{block indices} $q$ and $p$ \citep{takahashi2020convolutional}.
These indices are defined by the integer functions $\Delta k_{ij}$ and $\Delta l_{ij}$ 
(equations \ref{eq:Delta_kij} and \ref{eq:Delta_lij}), in terms of the indices $i$ 
of the observation points $(x_{i}, y_{i}, z_{0})$ and $j$ of the equivalent sources
$(x_{j}, y_{j}, z_{c})$.
For $x$-\textit{oriented grids} (Figure \ref{fig:regular-grids}), $Q = N_{y}$, $P = N_{x}$ 
and the block indices $q$ and $p$ are given by:
\begin{equation}
q \equiv q(i, j) = \Delta l_{ij}
\label{eq:q-x-oriented}
\end{equation}
and
\begin{equation}
p \equiv p(i, j) = \Delta k_{ij} \: ,
\label{eq:p-x-oriented}
\end{equation}
where $\Delta k_{ij}$ and $\Delta l_{ij}$ (equations \ref{eq:Delta_kij} and \ref{eq:Delta_lij}) 
are defined by integer functions $k(i)$, $k(j)$, $l(i)$ and $l(j)$ given by equations 
\ref{eq:k-x-oriented} and \ref{eq:l-x-oriented}.
For $y$-\textit{oriented grids} (Figure \ref{fig:regular-grids}), $Q = N_{x}$, $P = N_{y}$ and 
the block indices $q$ and $p$ are given by:
\begin{equation}
q \equiv q(i, j) = \Delta k_{ij}
\label{eq:q-y-oriented}
\end{equation}
and
\begin{equation}
p \equiv p(i, j) = \Delta l_{ij} \: ,
\label{eq:p-y-oriented}
\end{equation}
where $\Delta k_{ij}$ and $\Delta l_{ij}$ (equations \ref{eq:Delta_kij} and \ref{eq:Delta_lij}) 
are defined by integer functions $k(i)$, $k(j)$, $l(i)$ and $l(j)$ given by equations 
\ref{eq:k-y-oriented} and \ref{eq:l-y-oriented}.
Equations \ref{eq:q-x-oriented}--\ref{eq:p-y-oriented} show that $q$ varies from $-Q+1$
to $Q-1$ and $p$ from $-P+1$ to $P-1$, regardless of the grid orientation. They differ 
from those presented by \citet{takahashi2020convolutional} due to the absence of the module.

@@

Note that, for $x$-oriented grids, matrices $\Delta\mathbf{K}$ (equation \ref{eq:DK-matrix-x-oriented})
and $\Delta\mathbf{L}$ (equation \ref{eq:DL-matrix-x-oriented}) define the block indices
$p$ (equation \ref{eq:p-x-oriented}) and $q$ (equation \ref{eq:q-x-oriented}), respectively.
In this case, they are composed of $Q \times Q$ blocks with $P \times P$ elements each, where 
$Q = N_{y}$ and $P = N_{x}$. 
For $y$-oriented grids, matrices $\Delta\mathbf{K}$ (equation \ref{eq:DK-matrix-y-oriented})
and $\Delta\mathbf{L}$ (equation \ref{eq:DL-matrix-y-oriented}) define the block indices
$q$ (equation \ref{eq:q-y-oriented}) and $p$ (equation \ref{eq:p-y-oriented}), respectively.
In this case, they are also composed of $Q \times Q$ blocks with $P \times P$ elements each, 
but now $Q = N_{x}$ and $P = N_{y}$.
The examples shown by equations \ref{eq:DK-matrix-x-oriented}--\ref{eq:DL-matrix-y-oriented}
also illustrate that, regardless of grid orientation, (i) the block index $q$ is constant 
inside each block; (ii) blocks disposed along the same block diagonal are equal to each other; 
(iii) the block index $p$ is constant on each diagonal of a given block; 
(iv) elements of a given block located on the same diagonal are also equal do each other.
The results obtained with the small grid shown in Figure \ref{fig:regular-grids}
can be easily generalized for larger grids.
Based on the well-defined structure of block indices, we can define 
matrices $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ in a general form
\begin{equation}
\mathbf{A}_{\boldsymbol{\alpha\beta}} = \begin{bmatrix}
\mathbf{A}_{\boldsymbol{\alpha\beta}}^{0}   & \mathbf{A}_{\boldsymbol{\alpha\beta}}^{-1} & \cdots          & \mathbf{A}_{\boldsymbol{\alpha\beta}}^{-Q+1} \\
\mathbf{A}_{\boldsymbol{\alpha\beta}}^{1}   & \ddots          & \ddots          & \vdots           \\ 
\vdots           & \ddots          & \ddots          & \mathbf{A}_{\boldsymbol{\alpha\beta}}^{-1}   \\
\mathbf{A}_{\boldsymbol{\alpha\beta}}^{Q-1} & \cdots          & \mathbf{A}_{\boldsymbol{\alpha\beta}}^{1}  & \mathbf{A}_{\boldsymbol{\alpha\beta}}^{0}
\end{bmatrix}_{N \times N} \: ,
\label{eq:BTTB_A_alpha_beta}
\end{equation}
with blocks $\mathbf{A}_{\boldsymbol{\alpha\beta}}^{q}$, $q = -Q+1, \dots, Q-1$, given by
\begin{equation}
\mathbf{A}_{\boldsymbol{\alpha\beta}}^{q} = \begin{bmatrix}
a^{\alpha\beta}_{q0}   & a^{\alpha\beta}_{q(-1)} & \cdots  & a^{\alpha\beta}_{q(-P+1)} \\
a^{\alpha\beta}_{q1}   & \ddots     & \ddots  & \vdots       \\ 
\vdots      & \ddots     & \ddots  & a^{\alpha\beta}_{q(-1)}   \\
a^{\alpha\beta}_{q(P-1)} & \cdots     & a^{\alpha\beta}_{q1}  & a^{\alpha\beta}_{q0}
\end{bmatrix}_{P \times P} \: ,
\label{eq:Aq_block}
\end{equation}
formed by elements $a^{\alpha\beta}_{qp}$, $p = -P+1, \dots, P-1$.
This well-defined structure (equations \ref{eq:BTTB_A_alpha_beta} and \ref{eq:Aq_block}) 
of matrix components $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ 
(equation \ref{eq:A_expand}) is called Block-Toeplitz Toeplitz-Block (BTTB) 
\citep[e.g., ][ p. 67]{chan-jin2007}.

@@

Let $\mathbf{A}_{\boldsymbol{xz}}$ be a matrix component with elements
$a^{xz}_{ij}$ (equation \ref{eq:aij_alpha_beta}) defined by the second derivative
$h^{xz}_{ij}$ (equation \ref{eq:hxz_regular}). It can be easily verified from equations
\ref{eq:Delta_kij_symmetry}--\ref{eq:1_rij_symmetry} that $h^{xz}_{ij} = -h^{xz}_{ji}$.
As a consequence, $a^{xz}_{ij} = -a^{xz}_{ji}$, which means that 
\begin{equation}
\mathbf{A}_{\boldsymbol{xz}} = -\left( \mathbf{A}_{\boldsymbol{xz}} \right)^{\top}
\label{eq:Axz_symmetry}
\end{equation} 
for any grid orientation.
For $x$-oriented grids (Figure \ref{fig:regular-grids}), the block indices $q$ and $p$ are defined 
by equations \ref{eq:q-x-oriented} and 
\ref{eq:p-x-oriented} and $a^{xz}_{qp}$ can be rewritten as follows:
\begin{equation}
a^{xz}_{qp} = c_{m} \, \frac{\mu_{0}}{4\pi} 
\left( F_{x} u_{z} + F_{z} u_{x} \right) \frac{3 \left( p \, \Delta_{x} \right) \Delta_{z}}{r_{qp}^{5}}
\: ,
\label{eq:aqp_xz_x_oriented}
\end{equation}
with $\tfrac{1}{r_{qp}}$ defined by equation \ref{eq:1_rqp_x_oriented}.
In this case, we can see that
\begin{equation}
\mathbf{A}_{\boldsymbol{xz}}^{q} = \mathbf{A}_{\boldsymbol{xz}}^{(-q)}
\label{eq:Axz_q_external_block_symmetry_x_oriented}
\end{equation}
and 
\begin{equation}
\mathbf{A}_{\boldsymbol{xz}}^{q} = -\left( \mathbf{A}_{\boldsymbol{xz}}^{q} \right)^{\top} \: .
\label{eq:Axz_q_internal_block_symmetry_x_oriented}
\end{equation}
This structure is called \textit{symmetric-Block-Toeplitz skew symmetric-Toeplitz-Block} and is 
valid only for $x$-oriented grids.
For $y$-oriented grids (Figure \ref{fig:regular-grids}), the block indices $q$ and $p$ are 
defined by equations \ref{eq:q-y-oriented} and 
\ref{eq:p-y-oriented} and $a^{xz}_{qp}$ can be rewritten as follows:
\begin{equation}
a^{xz}_{qp} = c_{m} \, \frac{\mu_{0}}{4\pi} 
\left( F_{x} u_{z} + F_{z} u_{x} \right) \frac{3 \left( q \, \Delta_{x} \right) \Delta_{z}}{r_{qp}^{5}} \: ,
\label{eq:aqp_xz_y_oriented}
\end{equation}
with $\tfrac{1}{r_{qp}}$ defined by equation \ref{eq:1_rqp_y_oriented}.
Now, we conclude that
\begin{equation}
\mathbf{A}_{\boldsymbol{xz}}^{q} = -\mathbf{A}_{\boldsymbol{xz}}^{(-q)}
\label{eq:Axz_q_external_block_symmetry_y_oriented}
\end{equation}
and 
\begin{equation}
\mathbf{A}_{\boldsymbol{xz}}^{q} = \left( \mathbf{A}_{\boldsymbol{xz}}^{q} \right)^{\top} \: .
\label{eq:Axz_q_internal_block_symmetry_y_oriented}
\end{equation}
This structure is called \textit{skew symmetric-Block-Toeplitz symmetric-Toeplitz-Block} and is 
valid only for $y$-oriented grids.

The same reasoning can be followed to show that
\begin{equation}
\mathbf{A}_{\boldsymbol{yz}} = -\left( \mathbf{A}_{\boldsymbol{yz}} \right)^{\top}
\label{eq:Ayz_symmetry}
\end{equation} 
for any grid orientation. Besides, we can also show that
\begin{equation}
\mathbf{A}_{\boldsymbol{yz}}^{q} = -\mathbf{A}_{\boldsymbol{yz}}^{(-q)}
\label{eq:Ayz_q_external_block_symmetry_x_oriented}
\end{equation}
and 
\begin{equation}
\mathbf{A}_{\boldsymbol{yz}}^{q} = \left( \mathbf{A}_{\boldsymbol{yz}}^{q} \right)^{\top}
\label{eq:Ayz_q_internal_block_symmetry_x_oriented}
\end{equation}
for $x$-oriented grids (\textit{skew symmetric-Block-Toeplitz symmetric-Toeplitz-Block}), while
\begin{equation}
\mathbf{A}_{\boldsymbol{yz}}^{q} = \mathbf{A}_{\boldsymbol{yz}}^{(-q)}
\label{eq:Ayz_q_external_block_symmetry_y_oriented}
\end{equation}
and 
\begin{equation}
\mathbf{A}_{\boldsymbol{yz}}^{q} = -\left( \mathbf{A}_{\boldsymbol{yz}}^{q} \right)^{\top}
\label{eq:Ayz_q_internal_block_symmetry_y_oriented}
\end{equation}
for $y$-oriented grids (\textit{symmetric-Block-Toeplitz skew symmetric-Toeplitz-Block}).

@@

%======================================================================================
%\subsection{Standard Conjugate Gradient Least Squares (CGLS) method}
\subsection{Convolutional equivalent layer}
%======================================================================================

The computational cost associated with the classical method to estimate the parameter 
vector $\mathbf{p}$ by solving the linear system \ref{eq:normal-equations} can be very high 
or even prohibitive when dealing with large data sets. In these cases, a well-known alternative
is solving the normal equations (equation \ref{eq:normal-equations}) iteratively by 
using the \textit{standard Conjugate Gradient Least Squares (CGLS) method}:

@@

Setting a convergence criteria based on the minimum tolerance of the residuals is a good 
option to carry out this algorithm efficiently and still obtaining very good results. 
Another possibility is to set an invariance to the Euclidean norm of residuals between 
iterations, which would increase algorithm runtime, but with smaller residuals. 
We chose the first option, as we achieve satisfactory results.

Note that the standard CGLS solution (Algorithm \ref{al:std-cgls-algorithm}) requires 
neither inverse matrix nor matrix-matrix product. Instead, it only requires: one matrix-vector 
product out of the loop and two matrix-vector products per iteration (in steps 3 and 6). 
These products can be efficiently computed by using the 2D FFT, as a discrete convolution
(see Appendix A). \citet{takahashi2020convolutional} used this approach
to develop an efficient algorithm for gravity data processing. This modified approach in which
the standard CGLS method is modified to compute the matrix-vector products efficiently will be 
referenced throughout this work as the \textit{convolutional equivalent layer method}.

@@
